\documentclass[11pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{geometry}
\geometry{margin=2.5cm}

\title{Machine Learning on the OUALD Dataset:\\Classification and Regression}
\author{Andrei V\u{a}leanu \\ \texttt{andrei.valeanu03@stud.acs.upb.ro} \\ UPB -- Faculty of Automatic Control and Computers}
\date{2026}

\begin{document}
\maketitle

\begin{abstract}
This report summarises an introductory machine learning project on the Open University Learning Analytics Dataset (OUALD). Two supervised tasks are addressed: multi-class classification of the final course outcome (\texttt{final\_result}) and regression of the coursework score (\texttt{final\_coursework\_score}). The workflow includes exploratory data analysis, preprocessing, baseline and boosting models, and export of predictions for Moodle and Kaggle.
\end{abstract}

\section{Introduction}
The OUALD dataset contains demographic, engagement, and assessment features for Open University students. We predict the categorical outcome (Pass, Fail, Withdrawn, Distinction) and the numeric coursework score (0--100).

\section{Data and preprocessing}
\begin{itemize}
  \item \textbf{EDA:} class balance, correlations, chi-square and ANOVA tests.
  \item \textbf{Missing values:} median (numeric), mode (categorical).
  \item \textbf{Outliers:} IQR masking on \texttt{total\_clicks\_early}.
  \item \textbf{Features:} StandardScaler + one-hot (sklearn); native categories (CatBoost/XGBoost).
\end{itemize}

\section{Models}
\begin{itemize}
  \item Baselines: Decision Tree, Logistic Regression, Ridge, Lasso.
  \item Boosting: HistGradientBoosting, CatBoost (grid search), XGBoost blend.
  \item Kaggle: multi-seed ensemble + out-of-fold regression stacking for classification.
\end{itemize}

\section{Evaluation}
\begin{table}[h]
\centering
\begin{tabular}{lll}
\toprule
Task & Target & Metric \\
\midrule
Classification & \texttt{final\_result} & Accuracy \\
Regression & \texttt{final\_coursework\_score} & MSE \\
\bottomrule
\end{tabular}
\caption{Tasks and metrics.}
\end{table}

\section{Outputs}
Moodle: \texttt{CB\_OUALD\_predictii\_tema1.csv}. Kaggle: \texttt{kaggle\_TEMA1\_CB\_clasificare\_2026.csv}, \texttt{kaggle\_TEMA1\_CB\_regresie\_2026.csv}.

\section{Conclusion}
Boosting models outperform linear baselines on tabular OUALD data. CatBoost with categorical features and XGBoost blending provides the best results for the bonus competitions.

\begin{thebibliography}{9}
\bibitem{ouald} Open University Learning Analytics Dataset, \url{https://analyse.kmi.open.ac.uk/open-dataset/help/overview}
\bibitem{sklearn} scikit-learn, \url{https://scikit-learn.org/stable/}
\bibitem{catboost} CatBoost, \url{https://catboost.ai/en/docs/}
\bibitem{xgb} XGBoost, \url{https://xgboost.readthedocs.io/}
\end{thebibliography}

\end{document}
